Drawing Math

### 2022-05-13 paper/Paper.tex
fig

%% Custom Commands
\newcommand{\rationalPeriod}{{T^j}'}
\newcommand{\rationalPeriodTotal}{{T}'}
\newcommand{\statePeriod}{T^j}
\newcommand{\rationalAngleJ}[2]{{\theta^{#2}_{#1}}}
\newcommand{\rationalAngle}[1]{{\theta^j_{#1}}}
\newcommand{\rationalAngleNotime}{{\theta}^j}
\newcommand{\state}[2]{{s_{#1}}^{#2}}
\newcommand{\stateNotime}[1]{s^{#1}}
\newcommand{\totalPeriod}{T}
\newcommand{\totalState}{\pmb{s}_i}
\newcommand{\totalStateNotime}{\pmb{s}}
\newcommand{\comb}{Comb}
\newcommand{\combWithState}[1]{Comb\left(s^1_{#1}, s^2_{#1}, ..., s^k_{#1}\right)}
\newcommand{\inclIndic}[2]{{\mathrm{incl}_{#1}^{#2}}}
\newcommand{\numbToAngle}{\frac{2\pi}{b^j}}
\newcommand{\numbToAngleNoJ}[1]{\frac{2\pi}{b^{#1}}}
\newcommand{\combSingleTermCos}[1]{\prod_{j=1}^k \cos\left(\numbToAngle \cdot \theta^j\right) ^ {\inclIndic{#1}{j}}}
\newcommand{\digSumPeriodic}{\sigma^j}
\newcommand{\digSumPeriodicNoJ}{{\sigma}}
\newcommand{\totalOverRationalFrac}{\frac{\totalPeriod}{\rationalPeriodTotal}}
\newcommand{\changeTotalOneDim}{\Delta P_{0, T}^d}
\newcommand{\digit}{\mathrm{digit}}
\newcommand{\commonBase}{{\pmb{b}}}
\newcommand{\numbToCommon}{\frac{\commonBase}{b^j}}
\newcommand{\commonToAngle}{\frac{2\pi}{\commonBase}}
\newcommand{\sinOrCos}{\phi^j}
\newcommand{\combSingleTerm}[1]{\prod_{j=1}^k \sinOrCos\left(\numbToAngle \cdot \theta^j\right) ^ {\inclIndic{#1}{j}}}

@@

\section{Introduction}
A glum Pittsburgh day inspired the authors to take a random walk down YouTube's recommended
when they encountered the Numberphile video on ``Plotting Pi" \cite{NumberphileVid}.
In the video, Henderson and Brady introduce the idea of taking a Python Turtle and
deriving a series of updates to its position based off of various decimal sequences, 
some rational, some irrational. More specifically, they place an object in 2D, cartesian
space starting at $(0, 0)$ and ``facing" to the right. They then take a generating number, like $\pi$ or $35/99$, 
which gives a decimal sequence ($3.14...$ and $0.\bar{35}$ respectively). 
Then, at time step $i$, $i \geq 1$, they rotate the object by the $i$th digit divided by the base of
the decimal sequence. Then, the Turtle moves a constant distance in the direction which it faces.
For example, if the second digit is 4, the turtle is rotated by $\frac{4}{10}$ths of a circle
counterclockwise at time step 2 and then moves 10 units in its new direction.
% TODO: diagram, timestep 1 --> 2 --> 3

@@

\section{Background}
- Digital math
- Euler's formula
- I payed attention in some of my lectures

@@

\subsection{The problem}
Say we are given, $Comb$, hh$SU^j$, $p_0$, and $s^j_0$ for all $j \in [k]$.
Informally, the question is whether the turtle draws a ``closed" shape or not.\\
More formally, is there some period $T$ such that 
$$
  p_{i + \ell T} = p_{i}
$$
for $i, \ell \in \N$.
Then, note that if there exists a period $T$ such that 
$\Delta P_{i, i + \ell T} = 0$ for all $i, \ell \in N$, 
$p_{i + \ell T} = p_{i}$ and the turtle forms a closed shape.

\subsection{Specifying the task ahead of us}
For our case, we consider $Comb, SU_i^j$ to all be memoryless (i.e.\ there output
is uniquely determined by the current input). So, we can simplify the overall question.
If, $\pmb{s}_i = \pmb{s}_{i + \ell T}$ for some $T \in \N$ and all $i \in \N$, then 
$\Delta p_{i} = \Delta p_{i + \ell T}$. So then,
$\Delta P_{i, i + \ell T} = \Delta P_{i, i + \ell' T}$ for all $\ell, \ell' \in \N$.
Thus, $T$ is a period of the change in position. We can thus
break down our problem into two parts:
\begin{enumerate}
  \item Finding the period, $T$, of the state $\pmb{s}$.
  \item Checking whether $\Delta P_{i, i + T} = 0$.
\end{enumerate}

@@

\subsection{Property 1: Restricted Multinomials and Closure}
\label{prop:1}
Define $A_d = \set{j \mid j \in [k] \; \text{and}\; \inclIndic{d}{j} = 1}$, in other words,
$A_d$ is the set of rational parameters which are included in determining the position along the $d$th dimension.
Also, for function $f: \Z_\commonBase^{|A_d|} \rightarrow \Z_\commonBase$ and $\{a_1, a_2, ..., a_{|A_d|} \} = A_d$, we will denote
$$
  f\left(\digSumPeriodicNoJ^{a_1}, \digSumPeriodicNoJ^{a_2}, ..., \digSumPeriodicNoJ^{a_{A_d}}\right)
  = f\left(\digSumPeriodicNoJ\right).
$$

@@

If we were to then assume that $(\digSumPeriodicNoJ^1, ..., \digSumPeriodicNoJ^k)$ 
is uniformly and randomly draw from $\Z_\commonBase^k$, we then know that
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) = 0\right] < \frac{2^k}{\commonBase}
$$
by the Schwartz-Zippel Lemma \cite{Shwartz} \cite{Zip}.
So, this would leave us with
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) \neq 0\right] > 1 - \frac{2^k}{\commonBase}.
$$
In particular, this means that the probability of closure would be at least
$$
1 - \frac{2^k}{\commonBase}.
$$

@@

\subsection{Property 2: Root of Restricted Multinomial and Closure}
First note that
$$
\mathcal{M} = \{f : f(\pmb{x}) = \pm x_1 \pm x_2 ... \pm x_{|A_d|}\} = 
  \left\{\sum_{j \in |A_d|} (-1)^{\beta_j}\digSumPeriodic : \beta \in \set{0, 1}^{|A_d|} \right\}.
$$
Let $f_\beta \in \mathcal{M}$ then equal $\sum_{j \in |A_d|} (-1)^{\beta_j}\digSumPeriodic$.

Then, let
$$
  \mathcal{B} = \left\{\beta: f_\beta(\sigma) = 0\right\}
$$
and $\overline{\mathcal{B}} = \mathcal{M} \setminus \mathcal{B}$. Then, if for all
$\beta \in \mathcal{B}$,
$$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      I 
      \commonToAngle 
      \sum_{\ell = 0}^{q}
      \sum_{j \in A_d}
          \left(-1\right) ^ {\beta_{j}}
          \numbToCommon
          \mathrm{digit}(n, b, d, \ell)
  \right) = 0
$$
the Turtle will draw a closed shape. While this may seem arbitrary, there is a unique
and interesting geometric interpretation. 
Let rational sequence $a_1, a_2, ..., a_{\rationalPeriodTotal}$ be equal to
$
\forall \ell \in [\rationalPeriodTotal]
$,
$$
a_\ell = \sum_{j \in A_d}
\left(-1\right) ^ {\beta_{j}}
\numbToCommon
\mathrm{digit}(n, b, d, \ell).
$$
In other words, we are creating a ``common" rational sequence by summing and subtracting
our rational parameters in the least common multiple base $\commonBase$. Then,
if $$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
\exp\left(
    I 
    \commonToAngle 
    \sum_{\ell = 0}^{q}
      a_\ell
  \right) = 0
$$
the Turtle closes. If $r_i = \commonToAngle \sum_{\ell = 0}^{q} a_\ell$ for 
$i \in \set{0, ..., \rationalPeriodTotal - 1}$. Then $\exp(I \cdot r_i)$ can be thought of as some point around the 
unit circle. Also, note that $r_{i + 1} = r_i + \commonToAngle a_{i+1}$. In otherwords,
$r_i$ denotes some position around the complex unit circle where the subsequent updates
to position are given by a sum and subtraction of rational parameters. Then, if the center of
mass of these points around the unit circle is $0$, the Turtle draws a closed shape!
The authors are unsure as to the underlying intuition for why this is true.

@@

Over one common period for a all rational sequences, 
$\rationalPeriodTotal$,

@@

\section{Conclusion}

\section*{Acknowledgments}

>>>>

%% Custom Commands
\newcommand{\rationalPeriod}{{T^j}'}
\newcommand{\rationalPeriodTotal}{{T}'}
\newcommand{\statePeriod}{T^j}
\newcommand{\rationalAngleJ}[2]{{\theta^{#2}_{#1}}}
\newcommand{\rationalAngle}[1]{{\theta^j_{#1}}}
\newcommand{\rationalAngleNotime}{{\theta}^j}
\newcommand{\state}[2]{{s_{#1}}^{#2}}
\newcommand{\stateNotime}[1]{s^{#1}}
\newcommand{\totalPeriod}{T}
\newcommand{\totalState}{\pmb{s}_i}
\newcommand{\totalStateNotime}{\pmb{s}}
\newcommand{\comb}{Comb}
\newcommand{\combWithState}[1]{Comb\left(s^1_{#1}, s^2_{#1}, ..., s^k_{#1}\right)}
\newcommand{\inclIndic}[2]{{\mathrm{incl}_{#1}^{#2}}}
\newcommand{\numbToAngle}{\frac{2\pi}{b^j}}
\newcommand{\numbToAngleNoJ}[1]{\frac{2\pi}{b^{#1}}}
\newcommand{\combSingleTermCos}[1]{\prod_{j=1}^k \cos\left(\numbToAngle \cdot \theta^j\right) ^ {\inclIndic{#1}{j}}}
\newcommand{\digSumPeriodic}{\sigma^j}
\newcommand{\digSumPeriodicNoJ}{{\sigma}}
\newcommand{\totalOverRationalFrac}{\frac{\totalPeriod}{\rationalPeriodTotal}}
\newcommand{\changeTotalOneDim}{\Delta P_{0, T}^d}
\newcommand{\digit}{\mathrm{digit}}
\newcommand{\commonBase}{{\pmb{b}}}
\newcommand{\numbToCommon}{\frac{\commonBase}{b^j}}
\newcommand{\numbToCommonNoJ}[1]{\frac{\commonBase}{b^{#1}}}
\newcommand{\commonToAngle}{\frac{2\pi}{\commonBase}}
\newcommand{\sinOrCos}{\phi^j}
\newcommand{\combSingleTerm}[1]{\prod_{j=1}^k \sinOrCos\left(\numbToAngle \cdot \theta^j\right) ^ {\inclIndic{#1}{j}}}

@@

\section{Introduction}
A glum Pittsburgh day inspired the authors to take a random walk down YouTube's recommended
when they encountered the Numberphile video on ``Plotting Pi" \cite{NumberphileVid}.
In the video, Henderson and Brady introduce the idea of taking a Python Turtle and
deriving a series of updates to its position based off of various decimal sequences, 
some rational, some irrational. More specifically, they place an object in 2D, cartesian
space starting at $(0, 0)$ and ``facing" to the right. They then take a generating number, like $\pi$ or $35/99$, 
which gives a decimal sequence ($3.14...$ and $0.\bar{35}$ respectively). 
Then, at time step $i$, $i \geq 1$, they rotate the object by the $i$th digit divided by the base of
the decimal sequence. Then, the Turtle moves a constant distance in the direction which it faces.
For example, if the second digit is 4, the turtle is rotated by $\frac{4}{10}$ths of a circle
counterclockwise at time step 2 and then moves 10 units in its new direction.
\begin{figure}[ht]
  \centering
  \includegraphics[width=8cm]{fig/turtle2dtimestep.png}
  \caption{The turtle moving according to decimal sequence $0.223$ in base $10$.}
\end{figure}

@@

% \section{Background}
% - Digital math
% - Euler's formula
% - I payed attention in some of my lectures

@@

\subsection{The problem}
Say we are given, $Comb$, $SU^j$, $p_0$, and $s^j_0$ for all $j \in [k]$.
Informally, the question is whether the turtle draws a ``closed" shape or not.\\
More formally, is there some period $T$ such that 
$$
  p_{i + \ell T} = p_{i}
$$
for all $i, \ell \in \N$.
Then, note that if there exists a period $T$ such that 
$\Delta P_{i, i + \ell T} = 0$ for all $i, \ell \in N$, 
$p_{i + \ell T} = p_{i}$ and the turtle forms a closed shape.

\subsection{Specifying the task ahead of us}
For our case, we consider $Comb, SU_i^j$ to be uniquely determined by the current input.
So, we can simplify the overall question.
If, $\pmb{s}_i = \pmb{s}_{i + \ell T}$ for some $T \in \N$ and all $i \in \N$, then 
$\Delta p_{i} = \Delta p_{i + \ell T}$. So then,
$\Delta P_{i, i + \ell T} = \Delta P_{i, i + \ell' T}$ for all $\ell, \ell' \in \N$.
Thus, $T$ is a period of the change in position. We can thus
break down our problem into two parts:
\begin{enumerate}
  \item Finding the period, $T$, of the state $\pmb{s}$.
  \item Checking whether $\Delta P_{i, i + T} = 0$.
\end{enumerate}

@@

\subsection{Property 1: Restricted Multinomials and Closure}
\label{prop:1}
Define $A_d = \set{j \mid j \in [k] \; \text{and}\; \inclIndic{d}{j} = 1}$, in other words,
$A_d$ is the set of rational parameters which are included in determining the position along the $d$th dimension.
Also, for function $f: \Z_\commonBase^{|A_d|} \rightarrow \Z_\commonBase$ and $\{a_1, a_2, ..., a_{|A_d|} \} = A_d$, we will denote
$$
  f\left(\numbToCommonNoJ{a_1} \digSumPeriodicNoJ^{a_1}, \numbToCommonNoJ{a_2}\digSumPeriodicNoJ^{a_2}, ..., \numbToCommonNoJ{a_{A_d}}\digSumPeriodicNoJ^{a_{A_d}}\right)
  = f\left(\digSumPeriodicNoJ\right).
$$

@@

If we were to then assume that $(\digSumPeriodicNoJ^1, ..., \digSumPeriodicNoJ^k)$ 
is uniformly and randomly draw from $\Z_\commonBase^k$ and $\commonBase$ is a prime, we then know that
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) = 0\right] < \frac{2^k}{\commonBase}
$$
by the Schwartz-Zippel Lemma \cite{Shwartz} \cite{Zip}.
So, this would leave us with
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) \neq 0\right] > 1 - \frac{2^k}{\commonBase}.
$$
In particular, this means that the probability of closure would be at least
$$
1 - \frac{2^k}{\commonBase}.
$$

@@

\subsection{Property 2: Root of Restricted Multinomial and Closure}
First note that
$$
\mathcal{M} = \{f : f(\pmb{x}) = \pm x_1 \pm x_2 ... \pm x_{|A_d|}\} = 
  \left\{\sum_{j \in |A_d|} (-1)^{\beta_j} \numbToCommon \digSumPeriodic : \beta \in \set{0, 1}^{|A_d|} \right\}.
$$
Let $f_\beta \in \mathcal{M}$ then equal $\sum_{j \in |A_d|} (-1)^{\beta_j}\numbToCommon \digSumPeriodic$.

Then, let
$$
  \mathcal{B} = \left\{\beta: f_\beta(\sigma) = 0\right\}
$$
and $\overline{\mathcal{B}} = \mathcal{M} \setminus \mathcal{B}$. Then, if for all
$\beta \in \mathcal{B}$,
$$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      I 
      \commonToAngle 
      \sum_{\ell = 0}^{q}
      \sum_{j \in A_d}
          \left(-1\right) ^ {\beta_{j}}
          \numbToCommon
          \mathrm{digit}(n, b, d, \ell)
  \right) = 0
$$
the Turtle will draw a closed shape. While this may seem arbitrary, there is a unique
and interesting geometric interpretation. 
Let a rational decimal sequence $a_1, a_2, ..., a_{\rationalPeriodTotal}$ be equal to
$
\forall \ell \in [\rationalPeriodTotal]
$,
$$
a_\ell = \sum_{j \in A_d}
\left(-1\right) ^ {\beta_{j}}
\numbToCommon
\mathrm{digit}(n, b, d, \ell).
$$
In other words, we are creating a ``common" rational sequence by summing and subtracting
our rational parameters in the least common multiple base $\commonBase$. Then,
if $$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
\exp\left(
    I 
    \commonToAngle 
    \sum_{\ell = 0}^{q}
      a_\ell
  \right) = 0
$$
the Turtle closes.
If $r_i = \commonToAngle \sum_{\ell = 0}^{q} a_\ell$ for 
$i \in \set{0, ..., \rationalPeriodTotal - 1}$, then $\exp(I \cdot r_i)$ can be thought of as some point at angle $r_i$ around the 
complex unit circle. Also, note that $r_{i + 1} = r_i + \commonToAngle a_{i+1}$. In otherwords,
$r_i$ denotes some position around the complex unit circle where the subsequent updates
to position are given the $a_i$s, which are a sum and subtraction of rational parameters. Then, if the center of
mass of these points around the unit circle is $0$, the Turtle draws a closed shape!
The authors are unsure as to the underlying intuition for why this is true.

@@

@@

\section*{Acknowledgments}
The authors would like to thank their roommate Phu Nguyen for helping me work through the closure proofs
as well as inspiring me to pursue this project. The authors would also like to thank C.J. Argue
for assisting with the initial intuition and for providing encouragement. Moreover, the authors
would like to acknowledge Shiva Peri for doing a fantastic job putting together the artistic aspects
of the project as well as the website. Finally, the author would like to thank all their fantastic friends
from college who inspired and encouraged them to pursue their interests.
